---
title: "Extracting data for time slice analyses"
author: "Natalie Cooper and Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

Pipeline for getting data from discrete matrices.
===============

This code loads the data and extracts matrices and other files needed to run the analyses in the paper. Note that these analyses can take time, so we save the outputs to a `data/processed` folder and use these in later analyses so that the analyses can be replicated more quickly. Below we lay out the process in detail for one dataset, and then provide the code for the other datasets below this.

```{r}
set.seed(123)
```

Load packages
-----------------------------------

Make sure you work with `dispRity` 0.4 and TG's version of `Claddis` (much faster than GL's).

```{r, eval = TRUE, message = FALSE}
# The latest version of devtools and ape
library(devtools)
library(ape)

## The Thomas' Claddis development version
## (contains faster AncStateEstMatrix and MorphDistMatrixFast functions)
install_github("TGuillerme/Claddis")
library(Claddis)

## The latest version of dispRity 
install_github("TGuillerme/dispRity")
library(dispRity) 
```

Loading the data
-----------------------------------

This is based on a slug (here "Beck2014") and should work with different slugs, assuming the files are in the exact same format (by slug here we are referring to the fact that each of our files have the name of the dataset within them, so we can easily identify each one using regular expression matching). 

The analyses require a NEXUS matrix of characters, a csv file of first and last occurrence dates (FADLAD) and a phylogeny. The matrix and phylogeny need to be cleaned and manipulated - we do this below. The FADLAD can be read in as is so doesn't get read in until we need it for analyses in later scripts.

```{r}
## Setting up the slug name
slug <- "Beck2014"

## Loading the discrete morphological matrix
matrix <- ReadMorphNexus(paste0("../data/matrices/", slug, ".nex"))

## Loading the tree
tree <- read.nexus(paste0("../data/trees/", slug, ".tre"))

## Adding node labels
tree <- makeNodeLabel(tree, method = "number", prefix = "n")

## Adding a root age (max tree age date)
## Note that we need to redo this in the `02-time-slice-analyses` scripts
## as write.nexus/tree won't save root times
tree$root.time <- max(tree.age(tree)[, 1])
```

We'll also clean the data so that the species match in the tree and the matrix:

```{r}
## Cleaning the data 
cleaned_data <- clean.data(matrix$matrix, tree)

## Reassign the cleaned matrix and tree
matrix$matrix <- cleaned_data$data
tree <- cleaned_data$tree

## Removed taxa, i.e. those in the tree but not the matrix and vice versa
as.vector(na.omit(c(cleaned_data$dropped_tips, cleaned_data$dropped_rows)))

## Saving the tree
write.nexus(tree, file = paste0("../data/processed/tree.", slug ,".tre"))
```


Ancestral states reconstruction
-----------------------------------

Next we need to recreate the nodal states but excluding any ancestral state estimation with less than a 0.95 scaled Likelihood value. This allows us to discard any dodgy/ambiguous reconstructions (note that NAs are not a problem since we use Gower's distances later on).

This takes some time to run (could be a few hours)! Therefore this is set to `eval = FALSE`. The results are available in `data/processed`, or you can replace `FALSE` with `TRUE` to run this yourself.

```{r, eval = FALSE}
## Estimating the ancestral states (with a uncertainty threshold of 0.95)
## This is slow, so on future runs we recommend using the saved data
matrix_nodes <- AncStateEstMatrix(matrix, tree, estimate.allchars = TRUE,
                                  uncertainty.threshold = 0.95)

## Saving the matrix
saveRDS(matrix_nodes, file = paste0("../data/processed/ancestral_states.", slug ,".Rda"))
```

After this we can add the node states to the overall matrix:

```{r}
## Loading the matrix
## Not needed if running all the code at once
matrix_nodes <- readRDS(paste0("../data/processed/ancestral_states.", slug ,".Rda"))

## Adding the node names as row names
row.names(matrix_nodes) <- tree$node.label

## Combining the tips and the nodes in the same matrix
matrix$matrix <- rbind(matrix$matrix, matrix_nodes) 
```

Constructing the distance matrix
-----------------------------------

Again this takes some time to run (a couple of minutes). Therefore this is set to `eval = FALSE`. The results are available in `data/processed`, or you can replace `FALSE` with `TRUE` to run this yourself.

```{r, eval = FALSE}
## Calculating the Gower distance
matrix_dist <- MorphDistMatrixFast(matrix, distance = "Gower")

## Saving the matrix
saveRDS(matrix_dist, file = paste0("../data/processed/distance_matrix.", slug ,".Rda"))
```

One final issue is that to compute Gower distances we need all taxa to have some overlapping data. If not, Gower distances will be NA (Gower distance does not impute missing data like some other methods) and we can't estimate disparity. To avoid this problem we can remove the taxa with no overlaps using `TrimMorphDistMatrix` from `Claddis` on our distance matrix. Of course this then means we need to remove these from the tree as well. This then changes the node numbers which means we need to recalculate ancestral states. Because nothing is ever easy! Rather than going round in circles with this code, we just remove the species where this is a problem when we read in the trees (only an issue below for the Bapst and Brusatte datasets). Yes there's a more elegant way to do this, but time is tight and I'm pretty sure this project is trying to break us...

If you needed to do this for your own data, this is how you can quickly work out which taxa to remove.

```{r}
## Trimming the distance matrix
## trim <- TrimMorphDistMatrix(matrix_dist)
## matrix_dist <- trim$dist.matrix

## Which taxa were removed?
## (remove_taxa <- as.vector(trim$removed.taxa))
```

And that's it... we now have all the processed data for running the `dispRity` package properly: the distance matrices(that we will use to build the `morphospace`), and the cleaned tree (`tree`). Remember that for the `time.subsets` function, you can use the `inc.nodes` logical option to include nodes or not!

Running these extractions for other datasets
===================================================
Ideally we'd have written functions to do this, and/or a big old loop. But each dataset was slightly different so we've dealt with each individually.

Brusatte et al 2014 Current Biol
-----------

```{r}
slug <- "Brusatte2014"

## Loading the data and tree
matrix <- ReadMorphNexus(paste0("../data/matrices/", slug, ".nex"))
tree <- read.nexus(paste0("../data/trees/", slug, ".nex"))

# Replace zero branches with minimum branch length
tree$edge.length[which(tree$edge.length == 0)] <- 
  min(tree$edge.length[-which(tree$edge.length == 0)])

## Remove taxa with non overlapping data (see note about Gower distances above)
remove_Brusatte <- c("Shanag_ashile", "Atrociraptor_marshalli", "Proceratosaurus_bradleyi",     "Incisivosaurus_gauthieri", "Enigmosaurus", "Nanshiungosaurus_brevispinus", "Xixiasaurus", "Tsaagan_mangas", "Mirischia", "Pedopenna", "Suzhousaurus", "Juratyrant", "Vorona", "Bonapartenykus", "Teratophoneus", "Gobipteryx", "Songlingornis", "Liaoningornis_longidigitu", "Achillesaurus")

tree <- drop.tip(tree, remove_Brusatte)

# Tidy the tree and the data
tree <- makeNodeLabel(tree, method = "number", prefix = "n")
cleaned_data <- clean.data(matrix$matrix, tree)
matrix$matrix <- cleaned_data$data
tree <- cleaned_data$tree

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Remove the taxa not present in the tree but in the FADLAD
absent_in_tree <- clean.data(FADLAD, tree)$dropped_rows
if(!any(is.na(absent_in_tree))) {
    FADLAD <- FADLAD[-match(absent_in_tree, rownames(FADLAD)),]
}

## Getting the tree ages
tree_ages <- tree.age(tree)

## Scaling with the latest FADLAD
tree$root.time <- max(tree_ages[,1] + min(FADLAD[,2]))

## Removed taxa, i.e. those in the tree but not the matrix and vice versa
as.vector(na.omit(c(cleaned_data$dropped_tips, cleaned_data$dropped_rows)))

## Saving the tree
write.nexus(tree, file = paste0("../data/processed/tree.", slug ,".tre"))

## Estimating the ancestral states (with a uncertainty threshold of 0.95)
# matrix_nodes <- AncStateEstMatrix(matrix, tree, estimate.allchars = TRUE,
#                                  uncertainty.threshold = 0.95)
# saveRDS(matrix_nodes, file = paste0("../data/processed/ancestral_states.", slug ,".Rda"))

## Add ancestral states to nodes
matrix_nodes <- readRDS(paste0("../data/processed/ancestral_states.", slug ,".Rda"))
row.names(matrix_nodes) <- tree$node.label
matrix$matrix <- rbind(matrix$matrix, matrix_nodes) 

## Constructing the distance matrix
matrix_dist <- MorphDistMatrixFast(matrix, distance = "Gower")

## Trim taxa with non overlapping data
trim <- TrimMorphDistMatrix(matrix_dist)
matrix_dist <- trim$dist.matrix

## Which taxa were removed?
as.vector(trim$removed.taxa)

## Save
saveRDS(matrix_dist, file = paste0("../data/processed/distance_matrix.", slug ,".Rda"))
```

Bapst et al 2016 Biol Letters
-----------


```{r}
slug <- "Bapst2016"

## Loading the data and tree
matrix <- ReadMorphNexus(paste0("../data/matrices/", slug, ".nex"))
tree <- read.nexus(paste0("../data/trees/", slug, ".nex"))

## Replace zero branches with minimum branch length
tree$edge.length[which(tree$edge.length == 0)] <- 
  min(tree$edge.length[-which(tree$edge.length == 0)])

## Remove taxa with non overlapping data (see note about Gower distances above)
remove_Bapst <- c("Hagryphus_giganteus", "Atrociraptor_marshalli", "IGM100_1015_UndesDromaeosaurid", "Dromaeosaurus_albertensis", "Incisivosaurus_gauthieri", "Deinocheirus_mirificus", "Therizinosaurus_cheloniformis", "Anserimimus_planinychus", "Elmisaurus_rarus")

tree <- drop.tip(tree, remove_Bapst)

# Tidy the tree and the data
tree <- makeNodeLabel(tree, method = "number", prefix = "n")
cleaned_data <- clean.data(matrix$matrix, tree)
matrix$matrix <- cleaned_data$data
tree <- cleaned_data$tree

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Remove the taxa not present in the tree but in the FADLAD
absent_in_tree <- clean.data(FADLAD, tree)$dropped_rows
if(!any(is.na(absent_in_tree))) {
    FADLAD <- FADLAD[-match(absent_in_tree, rownames(FADLAD)),]
}

## Getting the tree ages
tree_ages <- tree.age(tree)

## Scaling with the latest FADLAD
tree$root.time <- max(tree_ages[,1] + min(FADLAD[,2]))

## Removed taxa, i.e. those in the tree but not the matrix and vice versa
as.vector(na.omit(c(cleaned_data$dropped_tips, cleaned_data$dropped_rows)))

## Saving the tree
write.nexus(tree, file = paste0("../data/processed/tree.", slug ,".tre"))

## Estimating the ancestral states (with a uncertainty threshold of 0.95)
# matrix_nodes <- AncStateEstMatrix(matrix, tree, estimate.allchars = TRUE,
#                                   uncertainty.threshold = 0.95)
# saveRDS(matrix_nodes, file = paste0("../data/processed/ancestral_states.", slug ,".Rda"))

## Add ancestral states to nodes
matrix_nodes <- readRDS(paste0("../data/processed/ancestral_states.", slug ,".Rda"))
row.names(matrix_nodes) <- tree$node.label
matrix$matrix <- rbind(matrix$matrix, matrix_nodes) 

## Constructing the distance matrix
matrix_dist <- MorphDistMatrixFast(matrix, distance = "Gower")
saveRDS(matrix_dist, file = paste0("../data/processed/distance_matrix.", slug ,".Rda"))
```

Wright 2017
-----------

```{r}
slug <- "Wright2017"

## Loading the data and tree
matrix <- ReadMorphNexus(paste0("../data/matrices/", slug, ".nex"))
tree <- read.nexus(paste0("../data/trees/", slug, ".tre"))

# Replace zero branches with minimum branch length
tree$edge.length[which(tree$edge.length == 0)] <- 
  min(tree$edge.length[-which(tree$edge.length == 0)])

# Note from Wright 2017: MCC tree. Note that to properly timescale the tree [e.g., in R] you must first divide the branch lengths by the corresponding clock rate, (= 0.03517385), and then set the root.time to 485.4]
# Divide branches by 0.03517385
tree$edge.length <- tree$edge.length / 0.03517385

# Add root time: 485.4
tree$root.time <- 485.4

# Tidy the tree and the data
tree <- makeNodeLabel(tree, method = "number", prefix = "n")
cleaned_data <- clean.data(matrix$matrix, tree)
matrix$matrix <- cleaned_data$data
tree <- cleaned_data$tree

## Removed taxa, i.e. those in the tree but not the matrix and vice versa
as.vector(na.omit(c(cleaned_data$dropped_tips, cleaned_data$dropped_rows)))

## Saving the tree
write.nexus(tree, file = paste0("../data/processed/tree.", slug ,".tre"))

## Estimating the ancestral states (with a uncertainty threshold of 0.95)
# matrix_nodes <- AncStateEstMatrix(matrix, tree, estimate.allchars = TRUE,
#                                   uncertainty.threshold = 0.95)
# saveRDS(matrix_nodes, file = paste0("../data/processed/ancestral_states.", slug ,".Rda"))

## Add ancestral states to nodes
matrix_nodes <- readRDS(paste0("../data/processed/ancestral_states.", slug ,".Rda"))
row.names(matrix_nodes) <- tree$node.label
matrix$matrix <- rbind(matrix$matrix, matrix_nodes) 

## Constructing the distance matrix
matrix_dist <- MorphDistMatrixFast(matrix, distance = "Gower")
saveRDS(matrix_dist, file = paste0("../data/processed/distance_matrix.", slug,".Rda"))

```