---
title: "Outputs from time slice analyses"
author: "Natalie Cooper and Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

This code runs the output analyses in the paper, i.e. creates tables and performs statistical tests.

```{r}
set.seed(123)
```

Load packages and functions
-----------------------------------

```{r, eval = TRUE, message = FALSE}
library(tidyverse)
library(broom)
library(dispRity)
source("../functions/stats_fun.R")
```

This script requires the outputs from `02-time-slice-analyses.Rmd`. This of course can all be written into functions but we ran out of time so here is some suboptimal but working code!

Testing for differences among the models is a little tricky, as we are interested in how disparity-through-time is influenced by the choice of time subsampling method, rather than the disparities estimated by each method, especially as these will influenced by the number of taxa (and/or nodes) included in each subsample. We therefore present two different, simple ways of comparing the models: paired Wilcoxon tests and identification of disparity peaks.

Paired Wilcoxon tests
----------------------

First, we use paired Wilcoxon tests to compare (i) unequal sized time bins to unevenly spaced time slices where the bins are stratigraphic periods (ages or epochs) and slices occur at the midpoint of the stratigraphic period; (ii) equal time bins to equally spaced time slices where time bin size, or the interval between slices is defined by the average *duration* of the stratigraphic period (age or epoch); and (iii) equal time bins to equally spaced time slices where the number of time bins, or number of slices is defined by the average *number* of stratigraphic periods (ages or epochs) in the time frame of interest.

Due to the uneven spread of taxa across phylogenies, some time bins will contain one or no species, meaning that we cannot estimate disparity for that time bin. We first, therefore, remove the time bins, and corresponding time slices, without disparity estimates. We then perform paired Wilcoxon tests, so that bins and slices for the same time period are being compared. Significant results suggest that there is a systematic difference in disparity values at each time point, depending on whether bins or slices are used.

```{r}
slugs <- c("Beck2014", "Brusatte2014", "Bapst2016", "Wright2017")
```

```{r}
# Loop through all four datasets
for(i in 1:4) {
  slug <- slugs[i]
  
#out1 <- readRDS(file = paste0("../outputs/age.", slug, ".Rda"))
out1 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

# i) statigraphy
strat <- out1$results$stratigraphy

# Work out how many disparity outputs per method there are
n_out <- 
  strat %>%
  filter(model == "acctran") %>%
  summarise(n = n())

# Get data into a format for wilcox tests
strat2 <- 
  strat %>% 
  # Add bin numbers so correct pairs can be compared  
  mutate(bin = rep(1:n_out$n, 7)) %>% 
  # Select only the required columns  
  select(bin, model, bs.median) %>%
  # Spread the models out by bin to make extracting NA bins easier
  spread(model, bs.median) %>%
  # Remove bins with NAs
  filter(complete.cases(.))

# Wilcox tests
w_acc <- wilcox.test(strat2$equalbins, strat2$acctran, paired = TRUE)
w_del <- wilcox.test(strat2$equalbins, strat2$deltran, paired = TRUE)
w_grad <- wilcox.test(strat2$equalbins, strat2$gradual.split, paired = TRUE)
w_prox <- wilcox.test(strat2$equalbins, strat2$proximity, paired = TRUE)
w_punc <- wilcox.test(strat2$equalbins, strat2$equal.split, paired = TRUE)
w_rand <- wilcox.test(strat2$equalbins, strat2$random, paired = TRUE)

# Save outputs
outputS <- rbind(tidy(w_acc), tidy(w_del), tidy(w_grad), tidy(w_prox), 
                tidy(w_punc), tidy(w_rand))
outputS$model <- c("acctran", "deltran", "gradual.split", "proximity", 
                        "equal.split", "random")
outputS$type <- rep("stratigraphy", 6)
outputS$period <- rep(strat$stratigraphy[1], 6)

## ii) duration

duration <- out1$results$duration

# Work out how many disparity outputs per method there are
n_out <- 
  duration %>%
  filter(model == "acctran") %>%
  summarise(n = n())

# Get data into a format for wilcox tests
duration2 <- 
  duration %>% 
  # Add bin numbers so correct pairs can be compared  
  mutate(bin = rep(1:n_out$n, 7)) %>% 
  # Select only the required columns  
  select(bin, model, bs.median) %>%
  # Spread the models out by bin to make extracting NA bins easier
  spread(model, bs.median) %>%
  # Remove bins with NAs
  filter(complete.cases(.))

# Wilcox tests
w_acc <- wilcox.test(duration2$equalbins, duration2$acctran, paired = TRUE)
w_del <- wilcox.test(duration2$equalbins, duration2$deltran, paired = TRUE)
w_grad <- wilcox.test(duration2$equalbins, duration2$gradual.split, paired = TRUE)
w_prox <- wilcox.test(duration2$equalbins, duration2$proximity, paired = TRUE)
w_punc <- wilcox.test(duration2$equalbins, duration2$equal.split, paired = TRUE)
w_rand <- wilcox.test(duration2$equalbins, duration2$random, paired = TRUE)

# Save outputs
outputD <- rbind(tidy(w_acc), tidy(w_del), tidy(w_grad), tidy(w_prox), 
                tidy(w_punc), tidy(w_rand))
outputD$model <- c("acctran", "deltran", "gradual.split", "proximity", 
                        "equal.split", "random")
outputD$type <- rep("duration", 6)
outputD$period <- rep(strat$stratigraphy[1], 6)

## iii) number

number <- out1$results$number

# Work out how many disparity outputs per method there are
n_out <- 
  number %>%
  filter(model == "acctran") %>%
  summarise(n = n())

# Get data into a format for wilcox tests
number2 <- 
  number %>% 
  # Add bin numbers so correct pairs can be compared  
  mutate(bin = rep(1:n_out$n, 7)) %>% 
  # Select only the required columns  
  select(bin, model, bs.median) %>%
  # Spread the models out by bin to make extracting NA bins easier
  spread(model, bs.median) %>%
  # Remove bins with NAs
  filter(complete.cases(.))

# Wilcox tests
w_acc <- wilcox.test(number2$equalbins, number2$acctran, paired = TRUE)
w_del <- wilcox.test(number2$equalbins, number2$deltran, paired = TRUE)
w_grad <- wilcox.test(number2$equalbins, number2$gradual.split, paired = TRUE)
w_prox <- wilcox.test(number2$equalbins, number2$proximity, paired = TRUE)
w_punc <- wilcox.test(number2$equalbins, number2$equal.split, paired = TRUE)
w_rand <- wilcox.test(number2$equalbins, number2$random, paired = TRUE)

# Save outputs
outputN <- rbind(tidy(w_acc), tidy(w_del), tidy(w_grad), tidy(w_prox), 
                tidy(w_punc), tidy(w_rand))
outputN$model <- c("acctran", "deltran", "gradual.split", "proximity", 
                        "equal.split", "random")
outputN$type <- rep("number", 6)
outputN$period <- rep(strat$stratigraphy[1], 6)

## And finally save all the outputs...

all_wilcox <- rbind(outputS, outputD, outputN)
all_wilcox <- select(all_wilcox, -c(method, alternative))
write.table(all_wilcox, file = paste0("../outputs/wilcox.", slug, ".csv"), quote = FALSE, 
          row.names = FALSE, col.names = FALSE, append = TRUE, sep = ",")
}
```

Repeat for epochs by uncommenting `out1 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))` and rerunning.

Disparity peaks
----------------------

```{r}
slugs <- c("Beck2014", "Brusatte2014", "Bapst2016", "Wright2017")
```

```{r, warning = FALSE}
# Loop through all four datasets
for(i in 1:4) {
  slug <- slugs[i]

## Read in the data
out1 <- readRDS(file = paste0("../outputs/age.", slug, ".Rda"))
out2 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

## Extract the maximum bootstrapped median disparity values for each model, with details of the ## peak bin/slice times.

## Getting the max/min for each age/epoch
age_results_max <- do.call(rbind, lapply(out1$objects, select.max.min.method, fun = max))
age_results_min <- do.call(rbind, lapply(out1$objects, select.max.min.method, fun = min))
epoch_results_max <- do.call(rbind, lapply(out1$objects, select.max.min.method, fun = max))
epoch_results_min <- do.call(rbind, lapply(out1$objects, select.max.min.method, fun = min))

## Combining into the nice table formats
age_results_max <- cbind(age_results_max, "stratigraphy" = rep("Age", nrow(age_results_max)),
                         "bin_type" = unlist(strsplit(rownames(age_results_max), split = "\\..*")))
age_results_min <- cbind(age_results_min, "stratigraphy" = rep("Age", nrow(age_results_min)),
                         "bin_type" = unlist(strsplit(rownames(age_results_min), split = "\\..*")))
epoch_results_max <- cbind(epoch_results_max, "stratigraphy" = rep("Epoch", nrow(epoch_results_max)),
                        "bin_type" = unlist(strsplit(rownames(epoch_results_max), split = "\\..*")))
epoch_results_min <- cbind(epoch_results_min, "stratigraphy" = rep("Epoch", nrow(epoch_results_min)),
                        "bin_type" = unlist(strsplit(rownames(epoch_results_min), split = "\\..*")))

## Combining the age and epoch results together
all_peaks_max <- rbind(age_results_max, epoch_results_max)
all_peaks_min <- rbind(age_results_min, epoch_results_min)

write.csv(all_peaks_max, file = paste0("../outputs/peaks.max.", slug, ".csv"), quote = FALSE, 
          row.names = FALSE)

write.csv(all_peaks_min, file = paste0("../outputs/peaks.min.", slug, ".csv"), quote = FALSE, 
          row.names = FALSE)
}
```


Extinction effect
------------------

Here we can test the effect of mass extinction on the different methods with the different models (for each dataset!).
The way to test the extinction effect is a referential wilcoxon test.#' @param as.list \code{logical}, whether to output the results as a list for \code{\link{test.dispRity}} (\code{TRUE}) or not (\code{FALSE} - default).

In other words, we compare the bins/slice just before the extinction event to the three ones just after (to account for an eventual lag effect).

Set the slug:

```{r}
slug <- "Beck2014" ; extinction <- 66.0
# slug <- "Brusatte2014"; extinction <- 66.0
# slug <- "Bapst2016"; extinction <- 66.0
# slug <- "Wright2017"; extinction <- 443.0
```

Read in the data

```{r}
out1 <- readRDS(file = paste0("../outputs/age.", slug, ".Rda"))
out2 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))
```

Measure differences in disparity:

```{r}
## Get the extinction times (ages)
ext_time_strat_age <- lapply(out1$objects$stratigraphy, extinction.subsets, extinction = extinction, lag = 3, as.list = TRUE)
ext_time_durat_age <- lapply(out1$objects$duration, extinction.subsets, extinction = extinction, lag = 3, as.list = TRUE)
ext_time_numbe_age <- lapply(out1$objects$number, extinction.subsets, extinction = extinction, lag = 3, as.list = TRUE)

## Get the extinction times (epochs)
ext_time_strat_epo <- lapply(out2$objects$stratigraphy, extinction.subsets, extinction = extinction, lag = 3, as.list = TRUE)
ext_time_durat_epo <- lapply(out2$objects$duration, extinction.subsets, extinction = extinction, lag = 3, as.list = TRUE)
ext_time_numbe_epo <- lapply(out2$objects$number, extinction.subsets, extinction = extinction, lag = 3, as.list = TRUE)

## Test the effect of extinctions
mapply.test <- function(data, times, test = wilcox.test, correction = "bonferroni") {
  test_run <- try(output <- test.dispRity(data, test = test, comparisons = times, correction = correction), silent = TRUE)
  if(class(test_run) == "try-error") {

    ## Try the test on less lags
    subsets_to_test <- unique(unlist(times))
    na_subsets <- summary(data)$obs[subsets_to_test]
    if(any(is.na(na_subsets))) {
      ## Remove the subsets with NA from the comparisons
      subsets_to_remove <- unlist(lapply(lapply(times, function(X) match(X, subsets_to_test[which(is.na(na_subsets))])), function(X) return(any(!is.na(X)))))

      ## Remove the subsets
      times_no_na <- times[!subsets_to_remove]
    } else {
      subsets_to_remove <- rep(TRUE, length(times))
    }

    ## Create an NA list
    comparison_names <- unlist(lapply(times, function(time, data) paste(names(data$subsets)[time], collapse = " : "), data))
    p_value <- statistic <- data.frame(rep(NA, 3), row.names = comparison_names)
    colnames(statistic) <- "statistic: W"
    colnames(p_value) <- "p.value"

    if(!all(subsets_to_remove)) {
      output <- test.dispRity(data, test = test, comparisons = times_no_na, correction = correction)

      ## Replacing the NAs
      statistic[!subsets_to_remove,1] <- output[[1]][,1]
      p_value[!subsets_to_remove,1] <- output[[2]][,1]
    }

    return(list(statistic, p_value))
  } else {
    return(output)
  }
}

## Applying the tests (ages)
test_strat_age <- mapply(mapply.test, out1$objects$stratigraphy, ext_time_strat_age, SIMPLIFY = FALSE)
test_durat_age <- mapply(mapply.test, out1$objects$duration, ext_time_durat_age, SIMPLIFY = FALSE)
test_numbe_age <- mapply(mapply.test, out1$objects$number, ext_time_numbe_age, SIMPLIFY = FALSE)

## Applying the tests (epochs)
test_strat_epo <- mapply(mapply.test, out2$objects$stratigraphy, ext_time_strat_epo, SIMPLIFY = FALSE)
test_durat_epo <- mapply(mapply.test, out2$objects$duration, ext_time_durat_epo, SIMPLIFY = FALSE)
test_numbe_epo <- mapply(mapply.test, out2$objects$number, ext_time_numbe_epo, SIMPLIFY = FALSE)

## Combining the tests into tables
wrap.up.results <- function(results, type, method) {
  results <- data.frame(matrix(unlist(results), ncol = 6, byrow = TRUE))
  ## Adding the models
  results <- cbind(c("bins", "acctran", "deltran", "random", "proximity", "equal.split", "gradual.split"),results)
  ## Adding the type and method
  results <- cbind(matrix(c(rep(type, 7), rep(method, 7)), ncol = 2), results)
  ## Adding the column names
  colnames(results) <- c("type", "method", "model", "e:1(W)", "e:2(W)", "e:3(W)", "e:1(p)", "e:2(p)", "e:3(p)")
  return(results)
}

## Combining the results
result <- wrap.up.results(test_strat_age, "age", "stratigraphy")
result <- rbind(result, wrap.up.results(test_durat_age, "age", "duration"))
result <- rbind(result, wrap.up.results(test_numbe_age, "age", "number"))
result <- rbind(result, wrap.up.results(test_strat_epo, "epoch", "stratigraphy"))
result <- rbind(result, wrap.up.results(test_durat_epo, "epoch", "duration"))
result <- rbind(result, wrap.up.results(test_numbe_epo, "epoch", "number"))

## Significant results
result <- cbind(result, apply(result[,7:9], 2, function(col) ifelse(col < 0.05, TRUE, FALSE)))

## Print the table
knitr::kable(result)

## Save the results
write.csv(result, file = paste0("../outputs/extinction.", slug, ".csv"), quote = FALSE, 
          row.names = FALSE)

```


Tables
----------

Finally, for the purposes of completeness here are the full results as tables.

Set the slug

```{r}
slug <- "Beck2014"
#slug <- "Brusatte2014"
#slug <- "Bapst2016"
#slug <- "Wright2017"
```

Read in the data

```{r}
out1 <- readRDS(file = paste0("../outputs/age.", slug, ".Rda"))
out2 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))
```

Make the tables

```{r, eval = TRUE, echo = FALSE}
# Ages
knitr::kable(out1$results$stratigraphy, caption = "Disparity per stratigraphic age: unequal time bins and non-equidistant time slices")
knitr::kable(out1$results$duration, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the median stratigraphic age duration")
knitr::kable(out1$results$number, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the number of stratigraphic ages")

# Epochs
knitr::kable(out2$results$stratigraphy, caption = "Disparity per stratigraphic epoch: unequal time bins and non-equidistant time slices")
knitr::kable(out2$results$duration, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the median stratigraphic epoch duration")
knitr::kable(out2$results$number, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the number of stratigraphic epochs")
```

You can rerun for the other datasets by changing the slug name at the start.